---
output:
  word_document: default
  html_document: default
---
# Basic Probability Distributions
## Contents

* The Normal Distribution
* The t Distribution
* The Binomial Distribution
* The Chi-Squared Distribution

### We look at some of the basic operations associated with probability distributions. There are a large number of probability distributions available, but we only look at a few. If you would like to know what distributions are available you can do a search using the command help.search("distribution").

### Here we give details about the commands associated with the normal distribution and briefly mention the commands for other distributions. The functions for different distributions are very similar where the differences are noted below.


### To get a full list of the distributions available in R you can use the following command:

```{r,include=FALSE}
# help(Distributions)

```

# Things to Remember: 

"d"	returns the height of the probability density function
"p"	returns the cumulative density function
"q"	returns the inverse cumulative density function (quantiles)
"r"	returns randomly generated numbers

# Normal Distributions

## Using Probablity density function

```{r}
dnorm(25, mean=50, sd=20)
# The Above code calculates 
# P(X=25) when X ~ N(50,20)

dnorm(25)
# When no mean is given, it takes the default values of mean=0 and standard deviation =1
```

## Cummilative Density function

```{r}
pnorm(20,56,10) 
# The Above code calculates 
# P(X<=20) when X ~ N(56,10)

pnorm(20,56,10,lower.tail=FALSE)
# The Above code calculates probablity of number larger than X
# P(X>20) when X ~ N(56,10)

```


## Inverse of Cummilative Density function
*  This return the number whose cumulative distribution matches the probability.
*  For example, if you have a normally distributed random variable with mean zero and standard deviation one, then if you give the function a probability it returns the associated Z-score:

```{r}
qnorm(0.5,1,2)
# Above code gives the number associated with probablity of 0.5 which has mean=1 and standard deviaion=2

```

##  Randomly generated numbers
* rnorm function which can generate random numbers whose distribution is normal. The argument that you give it is the number of random numbers that you want, and it has optional arguments to specify the mean and standard deviation

```{r}
rnorm(4,mean=3,sd=3)
# Generates 4 random number which are normally distributed with mean =3 and standard deviation =3

```

```{r}
# plotting the density function of a normal distribution: N(2, .25)
x <- seq(0, 4, 0.1) # creates a random set of numbers starting from 0 to 4 with breaks of 0.1
# Distribution of the x with probablity density function is plotted
plot(x, dnorm(x, 2, 0.5), type = "l")
```



#  T,binomial and chi squared distribution - please refer https://www.cyclismo.org/tutorial/R/probability.html#the-t-distribution

Here is a list of the functions that will generate a random sample from other common distributions: runif, rpois, rmvnorm, rnbinom, rbinom, rbeta, rchisq, rexp, rgamma, rlogis, rstab, rt, rgeom, rhyper, rwilcox, rweibull. Each function has its own set of parameter arguments. For example, the rpois function is the random number generator for the Poisson distribution and it has only the parameter argument lambda. The rbinom function is the random number generator for the binomial distribution and it takes two arguments: size and prob. The size argument specifies the number of Bernoulli trials and the prob argument specifies the probability of a success for each trial.
hypothesis tests for population means are done in R using the command "t.test".




One-sample hypothesis test

Let x represents a sample collected from a normal population with unknown mean and standard deviation. We want to test if the population mean is equal to 9, at significance level 5%.

The hypotheses are 
```{r}
x= c(6.2, 6.6, 7.1, 7.4, 7.6, 7.9, 8, 8.3, 8.4, 8.5, 8.6,
8.8, 8.8, 9.1, 9.2, 9.4, 9.4, 9.7, 9.9, 10.2, 10.4, 10.8,11.3, 11.9) # Entering the data

t.test(x-9,alternative="two.sided",conf.level=0.95)# Performing the t-test

```

Interpretation of the result

The P-value (0.3622) is greater than the significance level 5% (1-0.95), so we conclude that the null hypothesis that the mean of this population is 9 is plausible.

Two-sample hypothesis test

If we are interested in finding the confidence interval for the difference of two population means, the R-command "t.test" is also to be used.
The article "Measuring and Understanding the Aging of Kraft Insulating Paper in Power Transformers" (IIEE Electrical Insul. Mag., 1996:28-34), contained observations on degree of polymerization for paper specimens for which viscosity times concentration fell in a certain middle range 

We are interested in testing observations middle range and higher viscosity are from  populations with different means, at significance level 5%.

The hypotheses are 

```{r}
x=c(418,421,421,422,425,427,431,434,437,439,446,447,448,453,454,463,465)          # Entering the data into the R-workspace
y=c(429,430,430,431,36,437,440,441,445,446,447)

test2<-t.test(x,y,alternative="two.sided",mu=0,var.equal=F,conf.level=0.95)       # Performing a t-test procedure, containing a confidence level of 0.95
test2
```

Interpretation of the result
The P-value ( 0.3348) is greater than the significance level 5% (1-0.95), so we conclude that the null hypothesis that the population means are equal is plausible.



